---
title: "AN588 Homework 1"
subtitle: "Module 04: Working with Other Data Structures"
author: "Yongliang Cen"
date: "2025-01-31"
output: html_document
---

**Sherry's comment instructions:**

-   **I provided a summary of my suggestions in [red]{style="color: red;"}.**

-   **I put all code comments in #comment**

-   **Organization suggestions are already made**

-   **After each change I committed a change to your repo (hopefully you can see the comments)**

### CHALLENGE 1

First, create a vector of 1, word character strings comprising the first line of the novel Ulysses, by James Joyce: “Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.”

```{r}
# Load the stringr package
library(stringr)

# Create the string
my_line <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."

# Split word
split_word <- str_split(my_line, "\\s+",  simplify = TRUE)

# extract every third element
every_third_word <- split_word[seq(3, length(split_word), by = 3)]


every_third_word

```

### CHALLENGE 2

```{r}

# store the matrix
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)

#Given the matrix, m, above, extract the 2nd, 3rd, and 6th columns and assign them to the variable x
x <- m[, c(2, 3, 6)]

# Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x
x <- m[6:8, ]

# Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x
x <- m[2:6, 2:9]


```

### CHALLENGE 3

Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)

```{r}
a <- array(400:1, dim = c(5, 5, 4, 4))

a[1, 1, 1, 2]

a[2, 3, 2, ]

a[1:5, 1:5, 3, 3]

```

### Overwriting

```{r}
m[7, 1] <- 564
m[, 8] <- 2
m[2:5, 4:8] <- 1
m[2:5, 4:8] <- c(20, 19, 18, 17)
m[2:5, 4:8] <- matrix(data = c(20:1), nrow = 4, ncol = 5, byrow = TRUE)
m[, 8] <- c("a", "b")

```

### Lists and Data Frames

```{r}

s <- c("this", "is", "a", "vector", "of", "strings")  # here we have a vector
m <- matrix(data = 1:40, nrow = 5, ncol = 8)  # this is a matrix
b <- FALSE  #and this is a logical statement
l <- list(s, m, b)  # this puts them all together as a list
l

l[[2]]  #let's take another look at our matrix
l[[2]][2, 6]  #this will show us the entry at row 2, column 6 of the matrix in the second position of the list
l[1:2]  # here we're slicing our list so it just has the first and second elements
l[c(1, 3)]  # and here we're only taking the first and third elements
str(l)

names(l) <- c("string", "matrix", "logical")
names(l)

l$string  # this is calling the element named 'string' in list 'l'
l$matrix[3, 5]  # this is calling the entry in row 3, column 5 of the element named 'matrix' in list 'l'

```

### CHALLENGE 4

```{r}
# store the lowest family in vectors
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini  <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Tarsioidea  <- c("Tarsiidae")
Lorisoidea  <- c("Lorisidae", "Galagidae")
Lemuroidea  <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", "Daubentoniidae")

#group family into new vector that represent the higher family 
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- c("Platyrrhini", "Catarrhini")

Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- c("Anthropoidea", "Tarsioidea")

Strepsirhini <- list(Lorisoidea, Lemuroidea)
names(Strepsirhini) <- c("Lemuroidea", "Lorisoidea")

# combine all into "primates"
Primates <- list(Haplorhini, Strepsirhini)
names(Primates) <- c("Haplorhini", "Strepsirhini")

Primates

```

### Random People Data Frame

```{r}
# get directory for random_people.csv
getwd()

#read scv
df <- read.csv("~/Documents/GitHub/AN588-Homework1-briancen/random-people.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)

# only print select columns
df[, c(1, 3, 4, 11, 12)]

# check elements
str(df)

# returns a vector of data in column 3
df[, 3]


# returns a vector of data for column *name.last*
str(df$name.last)


# returns a data frame of data from column 3
df[3]

# returns a data frame of data for column *name.last*
df["name.last"]

# You can add rows (additional cases) or columns (additional variables) to a data frame using rbind() and cbind().
df <- cbind(df, id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16,
    17, 18, 19, 20))
df <- cbind(df, school = c("BU", "BU", "Harvard", "Harvard", "BU", "Boston College",
    "MIT", "BU", "BU", "UMASS Boston", "Harvard", "BU", "Boston College", "BU",
    "Harvard", "MIT", "Harvard", "BU", "UMASS Boston", "Harvard"))
df


# Alternatively, you can extend a data frame by adding a new variable directly using the $ operator, like this:
df$school <- c("BU", "BU", "Harvard", "Harvard", "BU", "Boston College", "MIT",
    "BU", "BU", "UMASS Boston", "Harvard", "BU", "Boston College", "BU", "Harvard",
    "MIT", "Harvard", "BU", "UMASS Boston", "Harvard")


```

### Filtering

```{r}
# Logical vectors may also be used to subset data frames. Here, we subset the data frame for only those rows where the variable school is “BU”. In this case, R evaluates the expression `df$school == “BU” and returns a logical vector equal in length to the number of rows in df. It then subsets df based on that vector, returning only rows that evaluate to TRUE.
new_df <- df[df$school == "BU", ]
new_df

# We can also choose to only return particular columns when we subset.
new_df <- df[df$school == "BU", c("name.last", "name.first", "school")]
new_df


```

### CHALLENGE 5

```{r}
numbers <- c(3, 0, 1,
          23, 1, 2,
          33, 1, 1,
          42, 0, 1,
          41, 0, 2)

# Create a 5×3 matrix, filled by row
m3 <- matrix(numbers, nrow = 5, ncol = 3, byrow = TRUE)

# View the matrix
print(m3)

# Convert matrix 'm3' to a data frame
df_m3 <- as.data.frame(m3)

# View the resulting data frame
print(df_m3)

# Coerce the 2nd column to logical
df_m3[, 2] <- as.logical(df_m3[, 2])


# Coerce the 3rd column factors
df_m3[, 3] <- as.factor(df_m3[, 3])

# show data type in df_m3
str(df_m3)

```
